\section{More Related Work}
\label{app:more_related_work}

\paragraph{Off-policy RLVR.}
Despite being framed as on-policy methods, practical RLVR implementations are
inherently off-policy: rollout data are reused across multiple gradient
updates~\citep{shao2024deepseekmathpushinglimitsmathematical,
huggingface_openr1_update3}; modern RL systems decouple rollout generation from
training for throughput~\citep{fu2025areal, slime_github,
noukhovitch2024asynchronous, zhong2025streamrl, he2025history}; and
training--inference mismatch from numerical precision differences can introduce
additional distribution shift~\citep{qi2025defeating}. To stabilize training
under off-policy conditions, prior work has proposed various strategies to
control policy updates, including ratio clipping~\citep{yu2025dapoopensourcellmreinforcement},
approximate trust regions~\citep{fu2025areal}, sequence-level
clipping~\citep{zheng2025group}, asymmetric clipping~\citep{roux2025tapered,
arnal2025asymmetric}, and gradient-preserving clipping~\citep{su2025klear,
chen2025minimax}. While these methods improve RLVR under moderate staleness, most focus on
relatively small staleness values and do not explicitly identify why naive
large-staleness GRPO fails. M2PO~\citep{zheng2025prosperitycollapsefaroffpolicy}
is the closest work to ours. It identifies excessive clipping under staleness and
the prosperity-before-collapse phenomenon, where training without a trust region
initially improves before eventually collapsing. M2PO addresses these issues by
constraining the second moment of importance weights at the batch level to
selectively mask high-variance tokens. In contrast, we diagnose the failure mode
of large-$\mu$ GRPO as localized negative-advantage suffix drift and stabilize
training with a simple trigger--scope filter, without requiring batch-level
second-moment statistics or iterative token masking. Moreover, our work
introduces an off-policy training paradigm that fully decou-
ples generation and optimization to fundamentally address
the efficiency bottleneck, an aspect M2PO does not explore.

\paragraph{RLVR Efficiency.}
RLVR systems improve efficiency through both synchronous/hybrid orchestration and
asynchronous execution. In synchronous pipelines, rollout generation and policy
optimization often alternate between inference and training engines. To make
frequent switching practical, systems may keep part of the training state
resident on GPU during rollout, reducing reload overhead but limiting memory
available for high-throughput inference. Hybrid orchestration frameworks such as
HybridFlow/verl improve scheduling across these engines~\citep{sheng2025hybridflow,
slime_github}, but frequent rollout refreshes and weight synchronization can
remain a bottleneck when the RL algorithm only tolerates low staleness. Asynchronous and streaming systems instead overlap rollout workers and trainers
through queues or streamed data transfer~\citep{fu2025areal, zhong2025streamrl,
han2025asyncflow, noukhovitch2024asynchronous}. These systems improve utilization,
but still benefit from staleness-robust objectives because larger tolerated
rollout lag reduces synchronization pressure. \ourmethod{} is complementary to
both directions: by stabilizing GRPO at $\mu=1024$, it reduces rollout refreshes
and weight synchronizations, enabling rollout generation to run as a standalone
phase that can fully use GPU memory for inference, while also permitting larger
lag in asynchronous pipelines.